Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-11-16 paper/main.tex
QA on Python/Cython

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-hashes.pdf}
    \caption{Tahoe-LAFS: Example of hashing operations.}
    \label{fig:tahoehashing}
    \emph{Figure based on
     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/CHK-hashes.svg}
     {CHK-hashes.svg from Tahoe-LAFS documentation}}
     %TODO: Footnote with full URL using footnotemark
\end{figure}

@@

\section{Python and Cython}
%What is it, how, why, alternatives?
% TODO: Finn en kilde på pythonting, om det trengs?
Python\footnote{\url{http://www.python.org}} is a high-level general-purpose
programming language. Python is also an interpreted language, which means that
python programs are compiled at runtime.  There exists multiple implementations
of Python, however the most common, CPython, is implemented in C. The fact that
Python is a high-level language usually means that there is performance penalty
in contrast with lower level languages, shuch as C. To remedy this problem it
is possible to write extensions to the language by using the official C-API.
This approach has been used by several cryptographic libraries for python such
as the official hashlib\footnote{\url{http://code.krypto.org/python/hashlib/}}
and pycryptopp\footnote{\url{http://tahoe-lafs.org/trac/pycryptopp}}.

Tahoe-\ac{LAFS} is written in python, but the SHA-3 candidate implementations found
in both the official NIST submissions and supercop are either in C or in
assembly. Therefore it is necessary to either implement the functions in python
or make extensions to python through the C-API. Where extensions should give
the best performance. % TODO: KILDE?

\subsection{Cython}
Cython\footnote{\url{http://www.cython.org}} is a language for writing Python
extensions in a language that closely resembles the Python language itself.
Basically what Cython does is to parse Cython files and generate C source files
that utilises the Python C-API. Afterwards the generated source files can be
compiled as if they were written directly in C. A small performance trade-off is
to be expected from this, due to the extra layer of abstraction.

@@

\paragraph{Blake32 \ac{SSSE3}} implementation also fail to wholly fulfill
criteria 4, however it does validate all test vectors that have a length 
in bits which is a multiple of 8. Since our python wrappers only allow, and
only need, to provide data in full bytes we decided to keep this
implementation.

@@

The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
This, along with the code being highly modular, emphasizes code readability.
With some experience using Python, it is not difficult to learn the Tahoe code
enough to understand what is happening when a file is uploaded or downloaded
with the system.

@@

As stated in Section \ref{sec:lengthextension}, none of the SHA-3 candidates is
vulnerable to the length-extension attack. Since we are interested in finding
out how Tahoe-\ac{LAFS} performs when utilizing the SHA-3 functions, we can
remove the lines that fixes the length-extension problems manually. In addition,
to try out one of the new implementations, we change line one in Listing
\ref{lst:hashutil}, and the result can be seen in Listing \ref{lst:hashutilmod}.

>>>>

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-hashes.pdf}
    \caption{Tahoe-LAFS: Example of hashing operations.}
    \label{fig:tahoehashing}
    \emph{Figure based on
     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/CHK-hashes.svg}
     {CHK-hashes.svg from Tahoe-LAFS documentation}}
\end{figure}

@@

\section{Python and Cython}
%What is it, how, why, alternatives?
% TODO: Finn en kilde på pythonting, om det trengs?
Python\footnote{\url{http://www.python.org/}} is a high-level general-purpose
programming language. Python is also an interpreted language, which means that
Python programs are compiled at runtime. There exists multiple implementations
of Python, however the most common, CPython, is implemented in C. The fact that
Python is a high-level language usually means that there are performance
penalties in contrast to lower level languages, such as C. To remedy this
problem it is possible to write extensions to the language by using the official
C-API. This approach has been used by several cryptographic libraries for
Python, such as pycryptopp\footnote{\url{http://tahoe-lafs.org/trac/pycryptopp}}
and the official hashlib\footnote{\url{http://code.krypto.org/python/hashlib/}}.

Tahoe-\ac{LAFS} is written in Python, but the SHA-3 candidate implementations found
in both the official NIST submissions and SUPERCOP are either in C or in
assembly. Therefore it is necessary to either implement the functions in Python
or make extensions to Python through the C-API, where extensions should give
the best performance.

\subsection{Cython}
Cython\footnote{\url{http://www.cython.org/}} is a language for writing Python
extensions in a language that closely resembles the Python language itself.
Basically what Cython does, is to parse Cython files and generate C source files
that utilises the Python C-API. Afterwards the generated source files can be
compiled as if they were written directly in C.

Direct calling of external C functions and methods is also supported, thus
making Cython an attractive way of wrapping external libraries written in C to
enable access from Python code.

Since the generation of C source files is an automated procedure, a small
performance trade-off is to be expected, due to the extra layer of abstraction.
The gain on the other hand, is significant in comparison to implementing
functions directly in Python.

@@

\paragraph{Blake32 \ac{SSSE3}} implementation also fail to wholly fulfill
criteria 4, however it does validate all test vectors that have a length 
in bits which is a multiple of 8. Since our Python wrappers only allow, and
only need, to provide data in full bytes we decided to keep this
implementation.

@@

The code of Tahoe-LAFS conforms to the style guide of Python, known as PEP 8.
This, along with the code being highly modular and consistent to its
architecture, emphasizes code readability.
With some experience using Python, it is not difficult to learn the Tahoe code
enough to understand what is happening when a file is uploaded or downloaded
with the system.

@@

As stated in Section \ref{sec:lengthextension}, none of the SHA-3 candidates is
vulnerable to the length-extension attack. Since we are interested in finding
out how Tahoe-\ac{LAFS} performs when utilizing the SHA-3 functions, we can
remove the lines that fixes the length-extension problems manually. In addition,
to try out one of the new implementations, we change line one in Listing
\ref{lst:hashutil}, and an example of this can be seen in Listing
\ref{lst:hashutilmod}.
